### 2016-01-14 04-evaluation.tex
Weak scaling + some strong scaling edits

\begin{table*}
  \centering
  \begin{tabular}{l r r l}
    Name            & \#Vertices &       \#Edges & Description \\
    \hline
    com-LiveJournal &  3,997,962 &    34,681,189 & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & Product co-purchasing network \\
    \hline
  \end{tabular}
  \label{tab-datasets}
  \caption{Summary of SNAP graph data sets used for evaluation.}
\end{table*}

@@

\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
  \caption{Strong scaling: Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes.}
  \label{fig-strong-scaling}
\end{figure*}

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability if its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-string-scaling}-a. As is clearly shown, the dominant phase of
the execution is update\_pi.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely, update\_pi and
deploy\_minibatch. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases
limiting their resource utilization.
%
The time spent in update\_beta remains relatively constant across cluster
sizes as it is an insignificant amount of work compared to the synchronization
overhead required to perform a collective MPI reduction operation to complete
it.

@@

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\columnwidth}
  \caption{Weak scaling: varying the number of compute nodes proportionally to
  the number of latent communities.}
  \label{fig-weak-scaling}
\end{figure}

\subsection{Effects of Number of Communities on Performance}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
  \caption{Effects of varying the number of communities on the algorithm's
  throughput.}
  \label{fig-throughput}
\end{figure}

\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
  \caption{FIXME}
  \label{fig-ppx-cpu}
\end{figure}

@@

\subsection{Convergence of Large Datasets}
\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx.eps, width=\textwidth}
  \caption{FIXME}
  \label{fig-ppx-cpu}
\end{figure*}

>>>>

\begin{table*}
  \centering
  \begin{tabular}{l r r l}
    Name            & \#Vertices &       \#Edges & Description \\
    \hline
    com-LiveJournal &  3,997,962 &    34,681,189 & Online blogging social network \\
    com-Friendster  & 65,608,366 & 1,806,067,135 & Online gaming social network \\
    com-Orkut       &  3,072,441 &   117,185,083 & Online social network \\
    com-Youtube     &  1,134,890 &     2,987,624 & Video-sharing social network \\
    com-DBLP        &    317,080 &     1,049,866 & Computer science bibliography collaboration network \\
    com-Amazon      &    334,863 &       925,872 & Product co-purchasing network \\
    \hline
  \end{tabular}
  \caption{Summary of SNAP graph data sets used for evaluation.}
  \label{tab-datasets}
\end{table*}

@@

\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-np-fixed-K.eps, width=\textwidth}
  \caption{(a) Execution time of 2048 algorithm iterations for the
  same problem size (com-Friendster, K=1024, M=16384 n=32) across different
  cluster sizes. (b) Speedup achieved for same experiments in (a) with respect to
  8-nodes}
  \label{fig-strong-scaling}
\end{figure*}

In order to evaluate the horizontal scalability of the distributed
implementation we tested the system's performance across different cluster
sizes while holding the problem size constant. For this study, we used the
com-Friendster graph as it contains the largest number of vertices and edges.
%
Figure~\ref{fig-strong-scaling}-a presents the execution time of 2048 algorithm
iterations across multiple cluster sizes. The x-axis starts from 8 worker nodes
as the data set is too large to fit into the collective memory of a smaller
cluster.
%
As shown in the figure, the execution
time steadily decreases by increasing the cluster size.
A deeper analysis of the individual execution phases of the algorithm provides
insights into the scalability if its building blocks.
%
In addition to the total execution time, Figure~\ref{fig-strong-scaling}-a
presents the cumulative time spent in individual computational phases across
iterations. Moreover, Figure~\ref{fig-strong-scaling}-b presents the speedup
achieved for the same experiments reported in
Figure~\ref{fig-strong-scaling}-a. As is clearly shown, the dominant phase of
the execution is update\_pi.
The reported total time for each cluster size is significantly less than the
sum of the execution times of the individual phases. This is due to the
overlapping execution of the two most expensive phases, namely, update\_pi and
deploy\_minibatch. Both of these phases initially gain significant speedup with the
addition of compute nodes. However, the speedup curve gradually slows down for
larger cluster sizes as the work granularity of each worker node decreases
limiting their resource utilization.
%
The time spent in update\_beta remains relatively constant across cluster sizes
as it performs an insignificant amount of work compared to the synchronization
overhead of a collective MPI reduction operation contained within it.

@@

\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-proportional-np.eps, width=\textwidth}
  \caption{(a) Average execution time per algorithm iteration varying
  the number of latent communities proportionally to the number of compute
  nodes. (b) The exact number of communities used for each data point in (a).}
  \label{fig-weak-scaling}
\end{figure*}

A study of a system's weak scaling aids in the assessment of the communication
and synchronization overheads associated with the management of large clusters.
Similarly, it can expose complex issues that hinder performance at scale such
as load imbalance. To fairly evaluate the algorithm's weak scaling behavior we
conducted several executions varying the cluster size and the number of latent
communities proportionally. This methodology ensures that each compute node
performs a relatively constant amount of computational work across the all
configurations. However, the number and size of messages exchanged between the
nodes would vary significantly. Figure~\ref{fig-weak-scaling}-a presents the
average execution
time per algorithm iteration across different cluster sizes. On the other hand,
Figure~\ref{fig-weak-scaling}-b reports the set number of communities for each
cluster configuration. The relative change in the average execution time per
iteration is insignificant even though the communication intensity increases
for larger cluster sizes. This observation verifies that the system's overall
overhead is minimal and scales well.

%################################################
\subsection{Effects of Number of Communities on Performance}

\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/sweep-over-K-fixed-np.eps, width=\columnwidth}
  \caption{Performance effect of varying the number of communities on the
    algorithm's execution time on 64-nodes.}
  \label{fig-k}
\end{figure}

\subsection{Horizontal vs. Vertical Scalability}
\begin{figure}[t] % [htb]
  \centering
  \epsfig{file=plots/hpc-cloud.eps, width=\columnwidth}
  \caption{FIXME}
  \label{fig-scale-up}
\end{figure}

@@

\subsection{Convergence of Large Datasets}
\begin{figure*}[t] % [htb]
  \centering
  \epsfig{file=plots/ppx.eps, width=\textwidth}
  \caption{FIXME}
  \label{fig-ppx}
\end{figure*}
